Multimatrices and Their Derivatives \\ \large An Impractical Guide to Global Conquest

### 2020-04-08 Math/book.tex
Proving some things about multiplicative inverses.

\DeclareMathOperator{\Dim}{Dim}
\DeclareMathOperator{\Ident}{I}
\DeclareMathOperator{\Tran}{Tran}
\DeclareMathOperator{\SoS}{\mathbb{S}}
\DeclareMathOperator{\Softmax}{Softmax}
\DeclareMathOperator{\Sum}{\Sigma}
\DeclareMathOperator{\remainder}{mod}
%\newcommand{\mmult}[1]{\stackrel{\times}{#1}}
%\newcommand{\mmult}[1]{\underset{#1}{\times}}
\newcommand{\mmult}[1]{\text{\raisebox{1ex}{$\underset{#1}{\times}$}}}
\newcommand{\dmult}[1]{\text{\raisebox{1ex}{$\underset{#1}{\,\bullet\,}$}}}

@@

That definition seems rather arbitrary but stick with me, I know what I'm
talking about. I've got a \sout{doctorate}
masters degree in \sout{mathematics} \sout{a real engineering field}, fine, software
engineering. Whatever. Just look at how the
square identity it relates to a multiplication operation.

@@

\begin{theorem}[Multiplicative Identity Implication]
\label{mm_ident_imp}
Similarly, if there is a matrix $X$ of shape $\vec{x} \oplus \vec{x}$ such that
for all matricies, $A$, for which the operation is defined,

@@

Then $X = \Ident^2(\vec{x})$.
\end{theorem}
\begin{proof}
TODO
\end{proof}

@@

Given a multimatrix, $X$, you might ask, ``What is the multiplicative inverse of $X$?''
And I would have to tell you that your question is not well formed.

\begin{align*}
  (M \mmult{m} X) \mmult{n} Y &= M \\
  |M| &= \vec{l} \oplus \vec{m} \\
  |X| &= \vec{m} \oplus \vec{x} \\
  |\vec{m}| &= m \\
  |M \mmult{m} X| &= \vec{l} \oplus \vec{x} \\
  |(M \mmult{m} X) \mmult{n} Y| &= |M| \\
  &= \vec{l} \oplus \vec{m} \\
  \therefore |Y| &= \vec{x} \oplus \vec{m} \\
  n &= |\vec{x}|
\end{align*}

By Thm. \ref{mm_associativity}, we can then say that,

\begin{align*}
  (M \mmult{m} X) \mmult{n} Y &= M \mmult{m} (X \mmult{n} Y) \\
  M &= M \mmult{m} (X \mmult{n} Y)
\end{align*}

Then by Thm. \ref{mm_ident_imp}, since $|X \mmult{n} Y| = \vec{m} \oplus \vec{m}$ and for all
$M$ where the operation is defined $M \mmult{m} (X \mmult{n} Y) = M$, it follows that
$X \mmult{n} Y = \Ident^2(\vec{m})$.

% TODO

>>>>

\DeclareMathOperator{\Dim}{Dim}
\DeclareMathOperator{\Ident}{I}
\DeclareMathOperator{\Tran}{Tran}
\DeclareMathOperator{\SoS}{\mathbb{S}}
\DeclareMathOperator{\Softmax}{Softmax}
\DeclareMathOperator{\Sum}{\Sigma}
\DeclareMathOperator{\remainder}{mod}
\DeclareMathOperator{\RMult}{RMult}
\DeclareMathOperator{\LMult}{LMult}
%\newcommand{\mmult}[1]{\stackrel{\times}{#1}}
%\newcommand{\mmult}[1]{\underset{#1}{\times}}
\newcommand{\mmult}[1]{\text{\raisebox{1ex}{$\underset{#1}{\times}$}}}
\newcommand{\dmult}[1]{\text{\raisebox{1ex}{$\underset{#1}{\,\bullet\,}$}}}

@@

That definition seems rather arbitrary but stick with me, I know what I'm
talking about. I've got a \sout{doctorate}
masters degree in \sout{mathematics} I mean \sout{a real engineering field}, fine, software
engineering. Whatever. Just look at how the
square identity it relates to a multiplication operation.

@@

There is an obvious corollary to that which you might consider an
alternate definition of $\Ident^2$.

\begin{corollary}[Multiplicative Identity Implication]
\label{mm_ident_imp}
If there is a matrix $X$ of shape $\vec{x} \oplus \vec{x}$ such that
for all matricies, $A$, for which the operation is defined,

@@

Then $X = \Ident^2(\vec{x})$.

Or if for all matricies, $B$, for which the operation is defined,

\[X \mmult{|\vec{x}|} B = B \]

Then, again, $X = \Ident^2(\vec{x})$.
\end{corollary}
\begin{proof}
Let $X$ be a matrix of shape $\vec{x} \oplus \vec{x}$. 
\begin{ppart}
If for all multimatrixes, $A$, on which $A \mmult{|\vec{x}|} X$ is defined, 
$A \mmult{|\vec{x}|} X = A$ this implies that, in the case of $A = \Ident^2(\vec{x})$,

\begin{align*}
  A \mmult{|\vec{x}|} X &= A \\
 \Ident^2(\vec{x}) \mmult{|\vec{x}|} X &= \Ident^2(\vec{x}) \\
  X &= \Ident^2(\vec{x}) 
\end{align*}
\end{ppart}
\begin{ppart}
If for all multimatrixes, $B$, on which $X \mmult{|\vec{x}|} B$ is defined, 
$X \mmult{|\vec{x}|} B = B$ this implies that, in the case of $B = \Ident^2(\vec{x})$,

\begin{align*}
  X \mmult{|\vec{x}|} B &= B \\
  X \mmult{|\vec{x}|} \Ident^2(\vec{x}) &= \Ident^2(\vec{x}) \\
  X &= \Ident^2(\vec{x}) 
\end{align*}
\end{ppart}
\end{proof}

@@

Given a multimatrix, $X$, you might ask, ``What is the multiplicative inverse of $X$?''
And I would have to tell you that your question is not well formed. Do you mean left multiplication
or right multiplication? Multiplication by what width? It's not clear which of these specifications
is even required: consider that multiplication is not communicative for regular 2d matricies but
those have simple multiplicative inverses (if an inverse exists).

So let's clarify what we can.

\begin{definition}[Multiplicative Functions]
Let us declare the function $\RMult[m,X](M)$ to be the function defined 
by the tuple,
\[ \left<m,X\right> \in \{  (\{0\} \cup \mathbb{N}) \times \mathbb{R}^{\vec{m} \oplus \vec{x}}  :  |\vec{m}| = m \} \]
such that,
\[ \RMult[m,X](M) : \mathbb{R}^{\vec{l} \oplus \vec{m}} \to \mathbb{R}^{\vec{l} \oplus \vec{x}} \]
\[ \RMult[m,X](M) = M \mmult{m} X \]

Let us also declare the function $\LMult[m,X](M)$ to be the function defined 
by the tuple,
\[ \left<m,X\right> \in \{  (\{0\} \cup \mathbb{N}) \times \mathbb{R}^{\vec{x} \oplus \vec{m}}  :  |\vec{m}| = m \} \]
such that,
\[ \LMult[m,X](M) : \mathbb{R}^{\vec{m} \oplus \vec{r}} \to \mathbb{R}^{\vec{x} \oplus \vec{r}} \]
\[ \LMult[m,X](M) = X \mmult{m} M \]

The the set of all these functions be $\RMult$ and $\LMult$ respectively.
\end{definition}

That will let us formally define a multiplicative inverse:

\begin{definition}[Multiplicative Inverse]
For any $\RMult[m,X]$ and $\RMult[n,Y]$ if, for all valid inputs of $\RMult[m,X]$, $M$,
\[ \RMult[n,Y] \circ \RMult[m,X] (M) = M \]
Then $\RMult[n,Y]$ is an inverse of $\RMult[m,X]$. In other words,
\[ \RMult[n,Y] \in \RMult[m,X]^{-1} \]

Similarly, for any $\LMult[m,X]$ and $\LMult[n,Y]$ if, for all valid inputs of $\LMult[m,X]$, $M$,
\[ \LMult[n,Y] \circ \LMult[m,X] (M) = M \]
Then $\LMult[n,Y]$ is an inverse of $\LMult[m,X]$. In other words,
\[ \LMult[n,Y] \in \LMult[m,X]^{-1} \]
\end{definition}

Note that those inverses are defined as sets of operations, not individual operations. 
You'll see why in a bit but to clear up any early confusion you should note that multimatrix
multiplicative inverses aren't unique. For example, $\RMult[2,\Ident^2(\left<2,2\right>)]$ has
two multiplicative inverse I can think of off the top of my head;
$\RMult[2,\Ident^2(\left<2,2\right)]$ or
$\RMult[1,\Ident^2(\left<2\right>)]$.

\begin{theorem}[Relation of Inverse to Identity]
If $\RMult[n,Y]$ is an inverse of $\RMult[m,X]$ where $|X| = \vec{m} \oplus \vec{x}$ and $|\vec{m}| = m$
then $X \mmult{n} Y = \Ident^2(\vec{m})$.
And if $\LMult[n,Y]$ is an inverse of $\LMult[m,X]$ where $|X| = \vec{x} \oplus \vec{m}$ and $|\vec{m}| = m$
then $Y \mmult{n} X = \Ident^2(\vec{m})$.
\end{theorem}
\begin{proof}
\begin{ppart}
Let $X$ be a multimatrix such that $|X| = \vec{m} \oplus \vec{x}$ and $|\vec{m}| = m$.
If $\RMult[n,Y]$ is the inverse of $\RMult[m,X]$ then, for all $M$,
\begin{align*}
  \RMult[n,Y] \circ \RMult[m,X] (M) &= M \\
  (M \mmult{m} X) \mmult{n} Y &= M
\end{align*}

This is also true in the case where $M = \Ident^2(\vec{m})$. In which case,
\begin{align*}
  (\Ident^2(\vec{m}) \mmult{m} X) \mmult{n} Y &= \Ident^2(\vec{m}) \\
  X \mmult{n} Y &= \Ident^2(\vec{m}) \\
\end{align*}
\end{ppart}
\begin{ppart}
Let $X$ be a multimatrix such that $|X| = \vec{x} \oplus \vec{m}$ and $|\vec{m}| = m$.
If $\LMult[n,Y]$ is the inverse of $\LMult[m,X]$ then, for all $M$,
\begin{align*}
  \LMult[n,Y] \circ \LMult[m,X] (M) &= M \\
  Y \mmult{n} (X \mmult{m} M) &= M
\end{align*}

This is also true in the case where $M = \Ident^2(\vec{m})$. In which case,
\begin{align*}
  Y \mmult{n} (X \mmult{m} \Ident^2(\vec{m})) &= \Ident^2(\vec{m}) \\
  Y \mmult{n} X &= \Ident^2(\vec{m})
\end{align*}
\end{ppart}
\end{proof}

\begin{theorem}[Shape of Inverses]
If $\RMult[n,Y]$ is an inverse of $\RMult[m,X]$ where $|X| = \vec{m} \oplus \vec{x}$ and
$|\vec{m}| = m$ then $n \le ||X||-m$.

Also, if $\LMult[n,Y]$ is an inverse of $\LMult[m,X]$ where $|X| = \vec{x} \oplus \vec{m}$ and
$|\vec{m}| = m$ then $n \le ||X||-m$.
\end{theorem}
\begin{proof}
\begin{ppart}
Let $\RMult[n,Y]$ be an inverse of $\RMult[m,X]$ where $|X| = \vec{m} \oplus \vec{x}$ and
$|\vec{m}| = m$.
Let $M$ be some valid input of $\RMult[m,X]$.
In other words, $|M| = \vec{l} \oplus \vec{m}$ for some $\vec{l}$.
So $|M \mmult{m} X| = \vec{l} \oplus \vec{x}$.
Since $\RMult[n,Y]$ is a valid operation on $M \mmult{m} X$, $n \le |\vec{l} \oplus \vec{x}|$.
And since there is no restriction on $\vec{l}$ it's only useful to say that $n \le |\vec{x}|$.
Given that $||X|| = |\vec{x}| + m$ it follows that $n \le ||X|| - m$.
\end{ppart}
\begin{ppart}
Let $\LMult[n,Y]$ be an inverse of $\LMult[m,X]$ where $|X| = \vec{x} \oplus \vec{m}$ and
$|\vec{m}| = m$.
Let $M$ be some valid input of $\LMult[m,X]$.
In other words, $|M| = \vec{m} \oplus \vec{r}$ for some $\vec{r}$.
So $|X \mmult{m} M| = \vec{x} \oplus \vec{r}$.
Since $\RMult[n,Y]$ is a valid operation on $X \mmult{m} M$, $n \le |\vec{x} \oplus \vec{r}|$.
And since there is no restriction on $\vec{r}$ it's only useful to say that $n \le |\vec{x}|$.
Given that $||X|| = |\vec{x}| + m$ it follows that $n \le ||X|| - m$.
\end{ppart}
\end{proof}

\begin{theorem}[Relation of Identity to Inverse]
If $X \mmult{n} Y = \Ident^2(\vec{m})$, $|\vec{m}| = m$, $|X| = \vec{m} \oplus \vec{x}$,
and $n \le ||X||-m$, that implies that $\RMult[n, Y]$ is an inverse of $\RMult[m, X]$.

Also,  if $Y \mmult{n} X = \Ident^2(\vec{m})$, $|\vec{m}| = m$, $|X| = \vec{x} \oplus \vec{m}$,
and $n \le ||X||-m$, that implies that $\LMult[n, Y]$ is an inverse of $\LMult[m, X]$.
\end{theorem}
\begin{proof}
\begin{ppart}
Let $X \mmult{n} Y = \Ident^2(\vec{m})$, $|\vec{m}| = m$, $|X| = \vec{m} \oplus \vec{x}$, and $n \le ||X||-m$.
We know that $n \le |\vec{x}|$ since $|\vec{x}| = ||X||-m$.
Then we can split up $\vec{x}$ into $\vec{x_l} \oplus \vec{x_r}$ where $|\vec{x_r}| = n$.
So $|Y| = \vec{x_r} \oplus \vec{y}$ for some $\vec{y}$ since $X \mmult{n} Y$ is defined.
We can then find the shape $|X \mmult{n} Y| = \vec{m} \oplus \vec{x_l} \oplus \vec{y}$.
Then for any $M$ such that $M \mmult{m} X$ is defined, i.e. any $M$ of shape $\vec{l} \oplus \vec{m}$ for some $\vec{l}$,
we can tell that $M \mmult{m} (X \mmult{n} Y)$ is also defined.
And since $n \le ||X||-m$ that implies $n+m \le ||X||$ so $M \mmult{m} (X \mmult{n} Y) = (M \mmult{m} X) \mmult{n} Y$
by Theorem \ref{mm_associativity}.
And if we just replace $X \mmult{n} Y$ with it's given value, then for all $M$ on which the $\RMult[m,X]$ is defined,
\begin{align*}
  (M \mmult{m} X) \mmult{n} Y &= M \mmult{m} (X \mmult{n} Y)  \\
  &= M \mmult{m} \Ident^2(\vec{m}) \\
  &= M \\
  \RMult[n,Y] \circ \RMult[m,X](M) &= M
\end{align*}
Therefore $\RMult[n,Y]$ is an inverse of $\RMult[m,X]$.
\end{ppart}
\begin{ppart}
Let $Y \mmult{n} X = \Ident^2(\vec{m})$, $|\vec{m}| = m$, $|X| = \vec{x} \oplus \vec{m}$, and $n \le ||X||-m$.
We know that $n \le |\vec{x}|$ since $|\vec{x}| = ||X||-m$.
Then we can split up $\vec{x}$ into $\vec{x_l} \oplus \vec{x_r}$ where $|\vec{x_l}| = n$.
So $|Y| = \vec{y} \oplus \vec{x_l}$ for some $\vec{y}$ since $Y \mmult{n} X$ is defined.
We can then find the shape $|Y \mmult{n} X| = \vec{y} \oplus \vec{x_r} \oplus \vec{m}$.
Then for any $M$ such that $X \mmult{m} M$ is defined, i.e. any $M$ of shape $\vec{m} \oplus \vec{r}$ for some $\vec{r}$,
we can tell that $(Y \mmult{n} X) \mmult{m} M$ is also defined.
And since $n \le ||X||-m$ that implies $n+m \le ||X||$ so $(Y \mmult{n} X) \mmult{m} M = Y \mmult{n} (X \mmult{m} M)$
by Theorem \ref{mm_associativity}.
And if we just replace $Y \mmult{n} X$ with it's given value, then for all $M$ on which the $\LMult[m,X]$ is defined,
\begin{align*}
  Y \mmult{n} (X \mmult{m} M) &= (Y \mmult{n} X) \mmult{m} M  \\
  &= \Ident^2(\vec{m}) \mmult{m} M\\
  &= M \\
  \LMult[n,Y] \circ \LMult[m,X](M) &= M
\end{align*}
Therefore $\LMult[n,Y]$ is an inverse of $\LMult[m,X]$.
\end{ppart}
\end{proof}

%TODO
